---
title: "Reproducible_Science_Assignment"
output: html_document
embed-resources: true 
---
```{r}
suppressMessages(library(tidyverse))
suppressMessages(library(palmerpenguins))
suppressMessages(library(ggplot2))
suppressMessages(library(janitor))
suppressMessages(library(here))
suppressMessages(library(tidyr))
renv::snapshot()
```


## QUESTION 01: Data Visualisation for Science Communication

*Create a figure using the Palmer Penguin dataset that is correct but badly communicates the data. **Do not make a boxplot**.*

*Use the following references to guide you:*

-   [*https://www.nature.com/articles/533452a*](https://www.nature.com/articles/533452a){.uri}
-   [*https://elifesciences.org/articles/16800*](https://elifesciences.org/articles/16800){.uri}

*Note: Focus on visual elements rather than writing misleading text on it.*

```{r}
names(penguins_raw)
```
```{r}
head(penguins_raw)
penguins_raw
```

```{r}
write.csv(penguins_raw, here("data", "penguins_raw.csv"))
```

```{r}
cleaning_penguins_function <- function(raw_data) {
  raw_data %>%
    remove_empty(c("rows", "cols")) %>% 
    clean_names() %>%
    drop_na(culmen_length_mm, body_mass_g) %>%
    mutate(species = case_when(
      species == "Adelie Penguin (Pygoscelis adeliae)" ~ "Adelie",
      species == "Chinstrap penguin (Pygoscelis antarctica)" ~ "Chinstrap",
      species == "Gentoo penguin (Pygoscelis papua)" ~ "Gentoo"
    )) 
}
```

```{r}
penguins_clean <- cleaning_penguins_function(penguins_raw)
```


```{r}
names(penguins_clean)
```

```{r}
write.csv(penguins_clean, here("data", "penguins_clean.csv"))
```

### a) Provide your figure here:

```{r}
# Creating a subset of the data
penguin_culmen_length <- penguins_clean %>% 
  select(species, culmen_length_mm)

head(penguin_culmen_length)
```


```{r bad figure code, echo=FALSE}
# There is no need to provide the code for your bad figure, just use echo=FALSE so the code is hidden. Make sure your figure is visible after you knit it. 
Fig1 <- penguin_culmen_length |>
  ggplot() +
  geom_histogram(aes(x = culmen_length_mm), binwidth = 0.05) +
  facet_wrap(~ species, ncol = 1) + xlab("Culmen length (mm)") +
  ylab("Count")
Fig1
```

### b) Write about how your design choices mislead the reader about the underlying data (200-300 words).

*Include references.*

------------------------------------------------------------------------

## QUESTION 2: Data Pipeline

*Write a data analysis pipeline in your .rmd RMarkdown file. You should be aiming to write a clear explanation of the steps, the figures visible, as well as clear code.*

### Introduction
We are interested in whether three different penguin species in the *Pygoscelis* genus have different body weights. 

```{r}
# Creating a subset of the data
penguin_body_mass <- penguins_clean %>% 
  select(species, body_mass_g)

head(penguin_body_mass)
```


```{r Data Exploration}
Fig2 <- penguin_body_mass |>
  ggplot(aes(x = species, y = body_mass_g, fill = species)) +
  geom_violin() +
  xlab("Penguin species") + ylab("Body mass (g)") +
  stat_summary(fun = mean, geom = "point", color = "black")
Fig2
```

### Hypothesis

### Statistical Methods

```{r Statistics}

# Make sure your code prints. 

```

### Results & Discussion

```{r Plotting Results}

# Make sure your code prints. 

```

### Conclusion

------------------------------------------------------------------------

## QUESTION 3: Open Science

### a) GitHub

*Upload your RProject you created for **Question 2** and any files and subfolders used to GitHub. Do not include any identifiers such as your name. Make sure your GitHub repo is public.*

*GitHub link:*

*You will be marked on your repo organisation and readability.*

### b) Share your repo with a partner, download, and try to run their data pipeline.

*Partner's GitHub link:*

*You **must** provide this so I can verify there is no plagiarism between you and your partner.*

### c) Reflect on your experience running their code. (300-500 words)

-   *What elements of your partner's code helped you to understand their data pipeline?*

-   *Did it run? Did you need to fix anything?*

-   *What suggestions would you make for improving their code to make it more understandable or reproducible, and why?*

-   *If you needed to alter your partner's figure using their code, do you think that would be easy or difficult, and why?*

### d) Reflect on your own code based on your experience with your partner's code and their review of yours. (300-500 words)

-   *What improvements did they suggest, and do you agree?*

-   *What did you learn about writing code for other people?*



